Disclosure of Product System Models in Life Cycle Assessment: Achieving Transparency and Privacy

### 2018-02-07 aggregation/agg-JIE.tex
Pauliuk comments
Headers for ERI submission
Supplementary stuffs
Multioutput case in SI
Streamlined Intro
Rewrite foreground study
Fig4 caption; disclosure + foreground work
Finishing body revision

\usepackage{tabularx}
\usepackage[round]{natbib}
\bibliographystyle{authordate_bk}

\renewcommand{\topfraction}{0.7}

@@

%\usepackage{lineno}
\let\showgrid\relax
\let\oldtimes\times
\let\times\cdot

@@

  \maketitle
  \emph{For submission to the Journal of Industrial Ecology; November 2017}

@@

  \doublespacing

@@

Objectives in critical review vary with the study and the reviewer, but they generally include consistency with the ISO 14044 standard, the validity of methods, appropriateness of data used, correctness of interpretations, and transparency of reporting \citep[Section 6.1]{iso14044}.  Although the object of critical review is typically a written report, the use of a structured disclosure of a PSM introduces the possibility that reviewers can inspect the model itself and even verify the computations and reported results.  This section discusses how a disclosure in the format described above can enable robust review of a product system model.  Note that, for purposes of critical review, the objective for data privacy can be omitted because the reviewer may be granted privileged access to confidential data.  We return to the data privacy requirement in the next section.

\subsection{Interpreting the Foreground}

The foreground of an LCA study is a collection of nodes that represent points in the inventory model where flows are exchanged.  The foreground matrix $A_f$ describes in precise terms what components are included in the product system and how they connect to one another.  The foreground is modelled as a weighted directed graph in which there is a 1:1 correspondence between nodes and product flows.  The foreground matrix $A_f$ reports the adjacency and weights of edges in the graph, and thus describes the relationships among these nodes.  These nodes typically represent unit processes, but they can also indicate exchanges across an implicit system boundary that encloses the foreground.  Generally, the foreground matrix $A_f$ can represent any process-flow model with a 1:1 correspondence between processes and flows, i.e. any allocated supply-and-use table, but some fundamental designs are common.  Some simple foreground models are shown in the supplementary materials.

@@

@@

 \singlespacing
 \bibliography{jabref}
}

\def\supporting{
  \renewcommand{\thefigure}{S\arabic{figure}}
  \renewcommand{\thetable}{S\arabic{table}}

@@

  \input{agg-foreground} % Af and graph/matrix equivalency

  \input{../manuscript/publication-example}

>>>>

%\lhead{\sffamily Disclosure of product system models}
%\rhead{\sffamily Kuczenski -- 2017}

\usepackage{tabularx}
\usepackage[round]{natbib}
\usepackage{siunitx}

\makeatletter
\let\oldsection\section
\renewcommand\section{\@startsection {section}{1}{\z@}%
                                   {-3.5ex \@plus -1ex \@minus -.2ex}%
                                   {2.3ex \@plus.2ex}%
                                   {\normalfont\Large\bfseries <Level 1> }*}
\renewcommand\subsection{\@startsection{subsection}{2}{\z@}%
                                     {-3.25ex\@plus -1ex \@minus -.2ex}%
                                     {1.5ex \@plus .2ex}%
                                     {\normalfont\large\bfseries <Level 2> }*}
\renewcommand\subsubsection{\@startsection{subsubsection}{3}{\z@}%
                                     {-3.25ex\@plus -1ex \@minus -.2ex}%
                                     {1.5ex \@plus .2ex}%
                                     {\normalfont\normalsize\bfseries <Level 3> }*}

\makeatother

\usepackage{endnotes}
\let\footnote=\endnote

\bibliographystyle{authordate_bk}

\renewcommand{\topfraction}{0.7}
\renewcommand{\bottomfraction}{0.7}

@@

\usepackage{lineno}
\let\showgrid\relax
\let\oldtimes\times
\let\times\cdot

@@

  \maketitle
  \emph{Submitted after revision to the Journal of Industrial Ecology; February 2018}

@@

  \doublespacing
  %\linenumbers

@@

Objectives in critical review vary with the study and the reviewer, but they generally include consistency with the ISO 14044 standard, the validity of methods, appropriateness of data used, correctness of interpretations, and transparency of reporting \citep[Section 6.1]{iso14044}.  Although the object of critical review is typically a written report, this level of review is not sufficient to ensure the correctness of quantitative results.
The use of a structured disclosure of a PSM introduces the possibility that reviewers can inspect the model itself and even verify the computations and reported results.  This section discusses how a disclosure in the format described above can enable robust review of a product system model. %\footnote{Although ISO 14071 makes explicit that reviewers are not required to review the system model, it is the opinion of both the author and at least one anonymous reviewer that the status quo is insufficient.}
Note that, for purposes of critical review, the objective for data privacy can be omitted because the reviewer may be granted privileged access to confidential data.  We return to the data privacy requirement in the next section.

\input{agg-foreground}

@@

%  \input{agg-community}

@@

\let\section\oldsection
    \section*{Acknowledgments}

Many thanks to the anonymous reviewers for significant improvement of the manuscript. This work was partially supported by the US National Science Foundation (CCF-1442966).

 \singlespacing

 \theendnotes

 \bibliography{jabref}
}

\def\supporting{
  \setcounter{figure}{0}
  \setcounter{table}{0}
  \renewcommand{\thefigure}{S\arabic{figure}}
  \renewcommand{\thetable}{S\arabic{table}}

@@

  \input{agg-preparing} % Af and graph/matrix equivalency

  \input{agg-multioutput}

  \input{publication-example}

### 2018-02-07 aggregation/agg-agg.tex
Fig4 caption; disclosure + foreground work
Finishing body revision

\subsubsection{Foreground or Foreground Fragment}

See Figure~\ref{fig:aggregation}b.  A foreground can be viewed as a linked set of unit processes.  The collection of processes themselves make up the foreground of the study, and so exchanges among them are recorded in $A_f$, and exchanges outside the foreground are represented in $A_d$ or $B_f$.  A complete foreground represents
the full disclosure of a model, and a reader can replicate and extend it independently.
%A {\red fragment} subset of the columns of $A_f$ may also be published  of the PSM can also be published as a complete foreground. %
Any co-production treatment, such as an allocation or system expansion, can only be fully expressed as a foreground fragment, because it intrinsically involves a collection of linked unit processes.

@@

See Figure~\ref{fig:aggregation}c. Any collection of foreground nodes can also be expressed in aggregated form by performing the computations in Eq.~\ref{eqn:agg}.  A publication of the aggregated foreground is identical to the publication of a unit process, except that the result was derived through aggregation.  As in the unit process case, the aggregated foreground has exactly one reference flow, the canonical functional unit $\tilde{\mathbf{y}}$.  The aggregated foreground still reports explicit links between the model and a background database and can thus be used to review data set selection.

>>>>

\subsubsection{Complete Foreground}

See Figure~\ref{fig:aggregation}b.  A foreground can be viewed as a linked set of unit processes.  The collection of processes themselves make up the foreground of the study, and so exchanges among them are recorded in $A_f$, and exchanges outside the foreground are represented in $A_d$ or $B_f$.  A complete foreground represents
the full disclosure of a model, and a reader can replicate and extend it independently.
%An excerpt from a larger study may also be published as a complete foreground. %
Any modeling technique used to treat a multi-functional process, such as an allocation or system expansion, can only be fully expressed as a foreground with at least as many columns as outputs.  For examples of multi-output processes represented as foreground models, please see the supporting information.  The example shown in Figure~\ref{fig:aggregation}b shows five nodes, including one cut-off flow

@@

See Figure~\ref{fig:aggregation}c. Any collection of foreground nodes can also be expressed in aggregated form by performing the computations in Eq.~\ref{eqn:agg}.  A disclosure of an aggregated foreground resembles the description of a unit process.  As in the unit process case, the aggregated foreground has exactly one reference flow, the canonical functional unit $\tilde{\mathbf{y}}$.  The aggregated foreground still reports explicit links between the model and a background database and can thus be used to review data set selection.

An aggregated foreground with cut-off flows resembles a multi-output process.  After aggregation, foreground flows that correspond to cut-offs must still be reported in a disclosure, while foreground flows interior to the model can be omitted.  This is done with the introduction of a new vector, $\tilde{\mathbf{a}}_f$, which includes the reference flow plus any entries from $\tilde{\mathbf{x}}$ that correspond to cut-off flows.

### 2018-02-07 aggregation/agg-conclusion.tex
Finishing body revision

This paper presents a possible solution by providing a mathematical formalization of the foreground of an LCA study and a functional specification for disclosing the PSM.  A data user or critical reviewer with access to the six disclosure elements, and also adequate access to background datasets and elementary flows used in the model, can reproduce the LCA computation in Eq.~\ref{eqn:study}, and ensure the reported result is correct.  Once equipped with the PSM, the reviewer or data user can potentially go much further.  Reviewers or data users who are able to reproduce the PSM from the disclosure can also modify it, adding or removing elements, altering dataset selections, and applying other impact assessment methods. Portions of the model which have been pre-aggregated can remain private in the disclosure. Further research is required to determine whether or under what conditions aggregation results can be reverse-engineered.

>>>>

This paper presents a possible solution by providing a mathematical formalization of the foreground of an LCA study and a functional specification for disclosing the PSM.  The disclosure framework also allows modelers to express precisely how multi-functionality is handled through allocation or system expansion.  A data user or critical reviewer with access to the six disclosure elements, and also adequate access to background datasets and elementary flows used in the model, can reproduce the LCA computation in Eq.~\ref{eqn:study}, and ensure the reported result is correct.  Once equipped with the PSM, the reviewer or data user can potentially go much further.  Practitioners who are able to reproduce the PSM from the disclosure can also modify it, adding or removing elements, altering dataset selections, applying other impact assessment methods, and considering alternative allocation methods. Portions of the model which have been pre-aggregated can remain private in the disclosure. Further research is required to determine whether or under what conditions aggregation results can be reverse-engineered.

### 2018-02-07 aggregation/agg-deps.tex
Disclosure card
Finishing body revision

\subsection{Dependencies, Emissions, and Cutoffs}

The foreground matrix only describes the structure of the PSM; it does not provide any information about inventory or impacts.  The inventory requirements and environmental burdens associated with the model are all described in the dependency matrix $A_d$ and the foreground emission matrix $B_f$.  The dependency matrix shows how foreground nodes depend on activities provided by processes in a background database; the emission matrix shows direct exchanges with the environment.  In both cases, the respective columns of $A_d$ and $B_f$ correspond with columns in $A_f$ and denote dependencies and/or emissions of the corresponding nodes.  Review of the dependencies of foreground nodes must include review of the selection of background processes to which they are connected (i.e. the identities of the columns of $B_x$ referred to in disclosure item~\ref{itm:bg}, the list of background flows). Likewise, the identities of the emissions (entries in disclosure item~\ref{itm:em}) must also be reviewed.

Interpretation of the contents of these matrices thus hinges on the contents of $A_f$.  If a column in $A_f$ is nonempty, then the corresponding node has exchanges with other nodes in the foreground.  If columns of $A_d$ and/or $B_f$ are empty, then the node is simply a pass-through process that serves to connect different foreground elements.  However, if $A_d$ and/or $B_f$ have nonzero entries in that column, then that indicates some transformation process or emission is occurring at that point in the foreground.

On the other hand, if a column in $A_f$ is empty, then that is an indication that the corresponding product flow is crossing the foreground system boundary.  In order for this flow to contribute any impacts, it must either be ``terminated'' in the dependency matrix, or represented as an emission.

Flows whose entire column consists of 0s throughout $A_f$, $A_d$, and $B_f$ are ``cut-offs'' that exit the model with zero burdens.  Specifying cut-off criteria and reporting of cut-off flows is an important part of system boundary definition.  Thus, reviewing cut-off flows (including evaluating their significance) is a crucial part of review that is facilitated by having a structured disclosure of the model.

>>>>

\subsection{Dependencies and Emissions}

The foreground matrix only describes the structure of the PSM; it does not provide any information about inventory or impacts.  The inventory requirements and environmental burdens associated with the model are all described in the dependency matrix $A_d$ and the foreground emission matrix $B_f$.  The dependency matrix shows how foreground nodes depend on activities provided by processes in a background database; the emission matrix shows direct exchanges with the environment.  In both cases, the respective columns of $A_d$ and $B_f$ correspond with columns in $A_f$ and denote dependencies and/or emissions of the corresponding nodes.  If columns of $A_d$ and/or $B_f$ are empty, then the node is simply a pass-through process that serves to connect different foreground elements.  However, if $A_d$ and/or $B_f$ have nonzero entries in that column, then that indicates some transformation process or emission is occurring at that point in the foreground.

Review of the dependencies of foreground nodes must include review of the selection of background processes to which they are connected (i.e. the identities of the columns of $B_x$ referred to in disclosure item~\ref{itm:bg}, the list of background flows). Likewise, the identities of the emissions (entries in disclosure item~\ref{itm:em}) must also be reviewed.

### 2018-02-07 aggregation/agg-disclosure.tex
Finishing body revision

\input{fig_disclosure}
Thus, the problem of how to disclose an LCA study design is reduced to the problem of how to communicate the contents of these submatrices accurately.  Recognizing that all three matrices are likely to be sparse (that is, that most of the entries are zero), the most efficient disclosure would be to only report non-zero entries.  A sparse matrix can be represented by a list of 3-tuples indicating (row, column, value) for each non-zero entry.  Thus, a formal disclosure of the PSM can be stated in six parts, including three lists of entities and three sets of data values:
\begin{enumerate}[label={\em d-\roman*}., ref={\em d-\roman*}]
\item\label{itm:fg} An ordered list of foreground nodes, beginning with the functional unit (rows/columns of $A_f$);
\item\label{itm:bg} An ordered list of background flows, each making reference to a particular background dataset (rows of $A_d$, mapping to columns of $B_x$);
\item\label{itm:em} An ordered list of exchanges with the environment (rows of $B_f$, mapping to entries in $\mathbf{e}$);
\item\label{itm:af} A set of 3-tuples for nonzero elements of $A_f$, in which the row and column are indices into Item~\ref{itm:fg};
\item\label{itm:ad} a set of 3-tuples for nonzero elements of $A_d$, in which the rows are indices into Item~\ref{itm:bg} and the columns are indices into Item~\ref{itm:fg};
\item\label{itm:bf} a set of 3-tuples for nonzero elements of $B_f$, in which the rows are indices into Item~\ref{itm:em} and the columns are indices into Item~\ref{itm:fg}.
\end{enumerate}
In both items~\ref{itm:bg} and~\ref{itm:em}, only rows containing non-zero entries need to be mentioned, since the index of each row is common information.  The contents of a disclosure are represented graphically in Fig.~\ref{fig:disclosure}., and two example disclosures drawn from reference databases can be found in the supplementary materials.

A disclosure containing these six elements satisfies most of the requirements established at the top of this section (the exception is ensuring the privacy of confidential data, which is treated later in the article):
\begin{itemize}
  \item The result is \textbf{computable} using Eq.~\ref{eqn:study};
  \item the computation is \textbf{complete} and \textbf{minimal} because all components are required and no superfluous components are included;
  \item The results are \textbf{reproducible} if the reader is able to accurately interpret the author's references in \ref{itm:bg} and \ref{itm:em};
  \item The disclosure includes only information for which the author can claim \textbf{authority};
  \item The disclosure includes only information for which the study is the \textbf{primary} source.  
\end{itemize}
As mentioned, the proper interpretation of the disclosure requires that the reader accurately interprets the author's references to background databases and elementary flows, including mutually consistent implementation of the LCIA vector $\mathbf{e}$.  These aspects are out of scope for the purposes of this paper, and we will assume that a shared understanding between author and reader can be reached.  However, the challenges of operationalizing this requirement are complex (see Discussion and Supplementary materials).  %Some implications on the information required for a successful disclosure are considered in the Discussion.

>>>>

\input{fig_disclosure}
Thus, the problem of how to disclose an LCA study design is reduced to the problem of how to communicate the contents of these submatrices accurately.  Recognizing that all three matrices are likely to be sparse (that is, that most of the entries are zero), the most efficient disclosure would be to only report non-zero entries.  A sparse matrix can be represented by a list of 3-tuples indicating (row, column, value) for each non-zero entry.  Thus, a formal disclosure of the PSM can be stated in six parts, including three lists of entities and three sets of data values:
\begin{enumerate}[label={\em d-\roman*}., ref={\em d-\roman*}]
\item\label{itm:fg} An ordered list of foreground nodes, beginning with the functional unit (rows/columns of $A_f$);
\item\label{itm:bg} An ordered list of background flows, each making reference to a particular background dataset (rows of $A_d$, mapping to columns of $B_x$);
\item\label{itm:em} An ordered list of exchanges with the environment (rows of $B_f$, mapping to entries in $\mathbf{e}$);
\item\label{itm:af} A set of 3-tuples for nonzero elements of $A_f$, in which the row and column are indices into Item~\ref{itm:fg};
\item\label{itm:ad} a set of 3-tuples for nonzero elements of $A_d$, in which the rows are indices into Item~\ref{itm:bg} and the columns are indices into Item~\ref{itm:fg};
\item\label{itm:bf} a set of 3-tuples for nonzero elements of $B_f$, in which the rows are indices into Item~\ref{itm:em} and the columns are indices into Item~\ref{itm:fg}.
\end{enumerate}
Each entity included in items \ref{itm:fg}--\ref{itm:em} describes a flow of some kind, and its unit of measure must be reported in the disclosure.
In both items~\ref{itm:bg} and~\ref{itm:em}, only rows containing non-zero entries need to be mentioned, since the index of each row is common information.  The contents of a disclosure are represented graphically in Fig.~\ref{fig:disclosure}., and two example disclosures drawn from reference databases can be found in the supporting information .

A disclosure containing these six elements satisfies most of the requirements established at the top of this section (the exception is ensuring the privacy of confidential data, which is treated later in the article):
\begin{itemize}
  \item The result is \textbf{computable} using Eq.~\ref{eqn:study};
  \item the computation is \textbf{complete} and \textbf{minimal} because all components are required and no superfluous components are included;
  \item The results are \textbf{reproducible} if the reader is able to accurately interpret the author's references in \ref{itm:bg} and \ref{itm:em};
  \item The disclosure includes only information for which the author can claim \textbf{authority};
  \item The disclosure includes all information specific to the study, or for which the study is the \textbf{primary} source, including re-implementations of already-published data. 
\end{itemize}
As mentioned, the proper interpretation of the disclosure requires that the reader accurately interprets the author's references to background databases and elementary flows, including mutually consistent implementation of the LCIA vector $\mathbf{e}$.  These aspects are out of scope for the purposes of this paper, and we will assume that a shared understanding between author and reader can be reached.  However, the challenges of operationalizing this requirement are complex (see the discussion and supporting information).  %Some implications on the information required for a successful disclosure are considered in the Discussion.

### 2018-02-07 aggregation/agg-foreground.tex
Finishing body revision

\section{Foreground Configurations}

Figure~\ref{fig:fragments} illustrates the structure of some basic foreground configurations.  Most product system models would include many of these elements mixed together.  A reviewer with access to items \ref{itm:fg} and \ref{itm:af} in the disclosure would be able to construct $A_f$ and automatically create a process-flow diagram, annotated with information about each node.

\input{fig_fragments}

The first (a) is a sequential model, in which each node requires one foreground input and generates one output. This model is equivalent to a ``gate to gate'' model.  Here the weights $k_i$ indicate the amount of the preceding reference flow that is required by the subsequent node.  Figure~\ref{fig:fragments}(b) shows an additive model, in which the outputs of several foreground nodes are added together, equivalent to a ``mixer'' or a horizontal average.  In this arrangement the weights %represent the relative weights of each input, and
should add up to a unit output of the reference node.  Finally, Figure~\ref{fig:fragments}(c) shows a foreground model with a cyclic dependency, where some of the reference output is consumed by another foreground node.

A typical PSM may contain multiple modules or fragments that are interconnected.  An example of foreground composed of several fragments is illustrated in Figure~\ref{fig:fragments}(d).  Here, the nodes labeled 1--5 represent one fragment, which generates the foreground's canonical reference flow $\tilde{y}$. This fragment requires two interior flows from separate fragments ($y_0$ and $y_1$), and has two unconnected flows (4 and 5).  The reference $y_0$ is supplied by a second fragment, made up of nodes 6-8.  The reference $y_1$ is supplied by another fragment made of only one node (9). The reference flow $y_1$ is consumed in two different places by the other fragments.

>>>>

\subsection{The Foreground System Boundary and Cut-off Flows}

The foreground of an LCA study is a collection of nodes that represent points in the inventory model where flows are exchanged.  The foreground matrix $A_f$ describes in precise terms what components are included in the product system and how they connect to one another.  The foreground is modelled as a weighted directed graph in which there is a 1:1 correspondence between nodes and product flows.  The foreground matrix $A_f$ reports the adjacency and weights of edges in the graph, and thus describes the relationships among these nodes.  Some simple foregrounds are described in the supporting information.

Generally, the direct requirements matrix can represent any process-flow model with a 1:1 correspondence between processes and flows.  Normally this is understood as a requirement to include only allocated single-output processes.  However, the foreground disclosure framework is also well suited to documenting how multi-output processes are transformed into single-output processes through either partitioning allocation or different forms of system expansion.  Different approaches for creating foregrounds to model multi-output process are presented in the supporting information.

An implicit ``foreground system boundary'' can be imagined to contain all the nodes in $A_f$.  Non-zero values in this matrix represent exchanges inside this boundary.  If a column in $A_f$ is nonempty, then the corresponding node has exchanges with other nodes in the foreground.  On the other hand, if a column in $A_f$ is empty, then that is an indication that the corresponding product flow is crossing the foreground system boundary.  In order for this flow to contribute any impacts, it must either be linked to a background process in the dependency matrix, or represented as an emission.

A flow whose entire column consists of 0s throughout $A_f$, $A_d$, and $B_f$ is a ``cut-off'' that exits the model with zero burdens.  Specifying cut-off criteria and reporting of cut-off flows is an important part of system boundary definition.  Thus, reviewing cut-off flows (including evaluating their significance) is a crucial part of review that is facilitated by having a structured disclosure of the model.

### 2018-02-07 aggregation/agg-lca-study.tex
Finishing body revision

The same result can be obtained by relocating the $\mathbf{y}$ vector into an augmented $A$ matrix, augmenting the $B$ matrix with a vector of zeros, and replacing the final demand with a vector $\tilde{\mathbf{y}}$ in which the first entry is 1 and all successive entries are 0:
\begin{equation}
  \begin{array}{cc}
s =& \mathbf{e}^T\times B\times\left(I-A\right)^{-1}\times\mathbf{y} \\
 =& \mathbf{e}^{T}\times\tilde{B}\times\left(I-\tilde{A}\right)^{-1}\times\tilde{\mathbf{y}}
  \end{array}
  \label{eqn:canonical}
\end{equation}
where
$\tilde{A} = \left[\begin{smallmatrix} 0 & \mathbf{0}^T \\ \mathbf{y} & A  \end{smallmatrix}\right]$,
$\tilde{B} = [ \mathbf{0},\, B ]$, and $\tilde{\mathbf{y}} =  [ 1,\, 0,\, 0 ,\,\ldots,\, 0]^{T}$.
It can easily be shown that Eq.~\ref{eqn:canonical} holds for any $\mathbf{y}$  (A proof is provided in the electronic supplementary information).  This is intuitively true because the $\tilde{\mathbf{y}}$ selects the first column of the augmented $\tilde{A}$ matrix as the ``final demand,'' but the ``direct requirements'' of that column are precisely the final demand from the prior, un-augmented $A$ matrix.

@@

\input{fig_foreground}
It is currently common practice for background database maintainers to pre-compute the life cycle inventory results for their databases.  In cases where the background database is used without modification, including the contents of $A$ in $\tilde{A}$ becomes unnecessary.
Using the notation $B_x = B \times (I - A)^{-1}$ to represent the pre-computed background LCI database, %% $\tilde{A}$ can be ``flattened'' so that it no longer includes inventories for background processes.
%% Eq.s~\ref{eqn:foreground} can be recast as:
%% \begin{equation}
%%  \tilde{A}_{flat} = \left[\begin{array}{cc} 
%% A_f &  0 \\
%% A_d & 0 
%%    \end{array}\right];\;\;\;  \tilde{B}_{flat} = \left[\begin{array}{cc} B_f & B_x \end{array}\right ]
%% \label{eqn:fg_bg}
%% \end{equation}
%% The proof of the equivalency of results derived from Eq.s~\ref{eqn:foreground} and \ref{eqn:fg_bg} is provided in the supplementary materials.  In this form, the only information remaining in $\tilde{A}_{flat}$ pertains to the study foreground.  
%
%% Using the ordered matrices, 
Eq.~\ref{eqn:leontief} can be written so that the foreground matrix inversion is  performed separately from the computationally costly background computation:
\begin{equation}
s = \mathbf{e}^T \times (B_f + B_x\times A_d) \times (I - A_f)^{-1} \times \tilde{\mathbf{y}}_f
\label{eqn:study}
\end{equation}
where $\tilde{\mathbf{y}}_f$ is a canonical functional unit vector whose length is the same as the rank of $A_f$. Eq.~\ref{eqn:study} is derived in the supplementary materials.
Eq.~\ref{eqn:study} is visualized in Figure~\ref{fig:foreground}, with the background technology matrix $A$ subsumed into $B_x$ and replaced by an identity matrix $I$.

>>>>

The same result can be obtained by relocating the $\mathbf{y}$ vector into an augmented $A$ matrix, augmenting the $B$ matrix with a vector of zeros, and replacing the final demand with a vector $\tilde{\mathbf{y}}$ in which the first entry is 1 and all successive entries are 0:
\begin{equation}
  \begin{array}{cc}
s =& \mathbf{e}^T\times B\times\left(I-A\right)^{-1}\times\mathbf{y} \\
 =& \mathbf{e}^{T}\times\tilde{B}\times\left(I-\tilde{A}\right)^{-1}\times\tilde{\mathbf{y}}
  \end{array}
  \label{eqn:canonical}
\end{equation}
where
$\tilde{A} = \left[\begin{smallmatrix} 0 & \mathbf{0}^T \\ \mathbf{y} & A  \end{smallmatrix}\right]$,
$\tilde{B} = [ \mathbf{0},\, B ]$, and $\tilde{\mathbf{y}} =  [ 1,\, 0,\, 0 ,\,\ldots,\, 0]^{T}$.
It can easily be shown that Eq.~\ref{eqn:canonical} holds for any $\mathbf{y}$  (A proof is provided in the electronic supporting information).  This is intuitively true because the $\tilde{\mathbf{y}}$ selects the first column of the augmented $\tilde{A}$ matrix as the ``final demand,'' but the ``direct requirements'' of that column are precisely the final demand from the prior, un-augmented $A$ matrix.

@@

\input{fig_foreground}
It is currently common practice for background database maintainers to pre-compute the life cycle inventory results for their databases.  In cases where the background database is used without modification, including the contents of $A$ in $\tilde{A}$ becomes unnecessary.
Using the notation $B_x = B \times (I - A)^{-1}$ to represent the pre-computed background LCI database, %% $\tilde{A}$ can be ``flattened'' so that it no longer includes inventories for background processes.
%% Eq.s~\ref{eqn:foreground} can be recast as:
%% \begin{equation}
%%  \tilde{A}_{flat} = \left[\begin{array}{cc} 
%% A_f &  0 \\
%% A_d & 0 
%%    \end{array}\right];\;\;\;  \tilde{B}_{flat} = \left[\begin{array}{cc} B_f & B_x \end{array}\right ]
%% \label{eqn:fg_bg}
%% \end{equation}
%% The proof of the equivalency of results derived from Eq.s~\ref{eqn:foreground} and \ref{eqn:fg_bg} is provided in the supplementary materials.  In this form, the only information remaining in $\tilde{A}_{flat}$ pertains to the study foreground.  
%
%% Using the ordered matrices, 
Eq.~\ref{eqn:leontief} can be written so that the foreground matrix inversion is  performed separately from the computationally costly background computation:
\begin{equation}
s = \mathbf{e}^T \times (B_f + B_x\times A_d) \times (I - A_f)^{-1} \times \tilde{\mathbf{y}}_f
\label{eqn:study}
\end{equation}
where $\tilde{\mathbf{y}}_f$ is a canonical functional unit vector whose length is the same as the rank of $A_f$. Eq.~\ref{eqn:study} is derived in the supporting information.
Eq.~\ref{eqn:study} is visualized in Figure~\ref{fig:foreground}, with the background technology matrix $A$ subsumed into $B_x$ and replaced by an identity matrix $I$.

### 2018-02-07 aggregation/agg-objective.tex
Further tweaking foreground derivation and proofs
Finishing body revision

The task of computing life cycle impact assessment (LCIA) results from a product system model can be described as an ``aggregation,'' in which the data in large, sparse models are combined through linear algebra operations into a small set of numeric scores.  Although several equivalent methods for computing aggregated LCI and/or LCIA results have been described in the literature \citep{Suh2005a, Peters_JLCA_2007}, the computational representation of and exchange of product system models, and the generation of aggregated results, remains software-specific.  
In the context of this paper, an ``LCA Study'' should be understood as any collection of inventory data under consideration as a group, from a single-operation unit process to a full ISO-style product LCA.

Reporting the results of an LCA study implies the existence of two parties with different information: the study author has full knowledge of the product system model; and the reader only has knowledge of what the author has disclosed.  Obviously, the level of detail of the disclosure will determine how useful it is to the reader.  In this context, the reader will want to know the answers to various review objectives.  
At the same time the author may have constraints that limit what can be disclosed.  The main goal of the framework is to give form to the disclosure so that the requirements of various review objectives, and restrictions associated with various disclosure contraints, can be evaluated clearly.

On the basis of transparency, the framework should meet the following objectives:
\begin{itemize}
\item \textbf{Computability}. The result should be computed using only the disclosure, plus resources available independently to the reader.  The computation should follow the consensus understanding of process LCA computation.
\item \textbf{Completeness / Minimality}.  The computation should use all of the information provided in the disclosure.
\item \textbf{Reproducibility}.  The result obtained by the reader should match the result provided by the author.
\end{itemize}
One key objective of disclosure is to inspect the work of the author.  Consequently, the disclosure should be made up of the author's own work product and should exclude, for instance, data that were drawn from a reference database or other independent source and used in an unmodified form.  The purpose of this is twofold: to focus the efforts of the reader on the work of the author, and to enable the reader to obtain and inspect the third-party data independently, thus ensuring its integrity.
\begin{itemize}
\item \textbf{Authority}.  The contents of the disclosure should authentically represent the outcomes of the author's labor in constructing the PSM.
\item \textbf{Primacy}.  The disclosure should exclude information that is made available by a third party.  (It must, however, still include previously-published data that were re-implemented by the author in the study.)
\end{itemize}
In addition, because of the ubiquity of confidential and sensitive information in LCA, any operational disclosure framework must also meet the following:
\begin{itemize}
\item \textbf{Privacy}.  The disclosure should protect information that is held in confidence by the author.
\end{itemize}

>>>>

The task of computing life cycle impact assessment (LCIA) results from a product system model can be described as an ``aggregation,'' in which the data in large, sparse models are combined through linear algebra operations into a small set of numeric scores.  Although several equivalent methods for computing aggregated LCI and/or LCIA results have been described in the literature \citep{Suh2005a, Peters_JLCA_2007}, the computational representation of and exchange of product system models, and the generation of aggregated results, remains software-specific.  
%In the context of this paper, an ``LCA Study'' should be understood as any collection of inventory data under consideration as a group, from a single-operation unit process to a full ISO-style product LCA.

Reporting the results of an LCA study implies the existence of two parties with different information: the study author has full knowledge of the product system model; and the reader only has knowledge of what the author has disclosed.  Obviously, the level of detail of the disclosure will determine how useful it is to the reader.  In this context, the reader will want to know the answers to various review objectives.  
At the same time the author may have constraints that limit what can be disclosed.  The main goal of the framework presented here is to give form to the disclosure so that the requirements of various review objectives, and restrictions associated with various disclosure contraints, can be evaluated clearly.

On the basis of transparency, the framework for disclosure of the PSM should meet the following objectives:
\begin{itemize}
\item \textbf{Computability}. The result should be computed using only the disclosure, plus resources available independently to the reader.  The computation should follow the consensus understanding of process LCA computation.
\item \textbf{Completeness / Minimality}.  The computation should use all of the information provided in the disclosure.
\item \textbf{Reproducibility}.  The result obtained by the reader should match the result provided by the author.
\end{itemize}
One key objective of disclosure is to inspect the work of the author.  Consequently, the disclosure should be made up of the author's own work product and should exclude, for instance, data that were drawn from a reference database or other independent source and used in an unmodified form.  The purpose of this is twofold: to focus the efforts of the reader on the work of the author, and to enable the reader to obtain and inspect the third-party data independently, thus ensuring its integrity.
\begin{itemize}
\item \textbf{Authority}.  The contents of the disclosure should authentically represent the outcomes of the author's labor in constructing the PSM.
\item \textbf{Primacy}.  The disclosure should exclude information that is made available by a third party.  (It must, however, still include previously-published data that were re-implemented by the author in the study.)
\end{itemize}
In addition, because of the ubiquity of confidential and sensitive information in LCA, any operational disclosure framework must also meet the following:
\begin{itemize}
\item \textbf{Privacy}.  The disclosure should protect information that is held in confidence by the author.
\end{itemize}

### 2018-02-07 aggregation/agg-review.tex
Finishing body revision

\subsection{Comparative Review: Formalizing Scope and System Boundary}

The disclosure of a PSM contains six parts, including three lists and three tables of sparse matrix data.  The graph structure encoded in the foreground matrices can be automatically rendered as a process-flow diagram, supplementing or supplanting the hand-crafted system diagrams that currently accompany ISO reports. While the numeric data are required to critically evaluate and reproduce the study results, substantial information about the study design can be found in the list of background dependencies (\ref{itm:bg}) and emissions (\ref{itm:em}).  These lists report all the distinct sources of environmental impacts in the model.  The cutoff flows, which are foreground nodes that have no nonzero entries in $A_f$, $A_d$, or $B_f$, report on the other hand what aspects were explicitly excluded from the model.

>>>>

\subsection{Comparative Review: Formalizing Model Scope}

The disclosure of a PSM contains six parts, including three lists and three tables of sparse matrix data.  The graph structure encoded in the foreground matrices can be automatically rendered as a process-flow diagram, supplementing or supplanting the hand-crafted system diagrams that currently accompany ISO reports. While the numeric data are required to critically evaluate and reproduce the study results, substantial information about the study design can be found in the list of background dependencies (\ref{itm:bg}) and emissions (\ref{itm:em}).  These lists report all the distinct sources of environmental impacts in the model.  The cut-off flows, which are foreground nodes that have no nonzero entries in $A_f$, $A_d$, or $B_f$, report on the other hand what aspects were explicitly excluded from the model.
